### 2022-06-15 paper.tex
add Ana's comments

While this simple procedure is straight-forward and produces satisfactory results, there is room for improvement. Here, we propose three improvements that, as we show afterwards, considerably increase the information content of the processed data. We will first explain the proposed improvement in a qualitative manner, while mathematical details are given later in the Methods section. Then, we will demonstrate the improvement using benchmarks.

@@

We also presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms. The package also offers functionality for all other steps of data preprocessing that are needed to get from the output of methylation callers such as Bismark and methylpy to a cell\texttimes region matrix that is suitable as input for standard analysis tools that work in a reduced-dimensionality setting, thus allowing the user to use, e.g., well established tools from scRNA-seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses etc. By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis.

@@

We write $x_{ij}$ for the methylation status of CpG $i$ in cell $j$. The index $i$ runs over all CpG positions present in the genome, the index $j$ over all cells in the assay. We write $x_{ij}=0$ if position $i$ was found to be unmethylated in cell $j$ by bisulfite sequencing, $x_{ij}=1$ if it was methylated, and $x_{ij}=\text{NA}$ if position $i$ was not covered by reads from cell $j$ and the methylation status is therefore not available ("NA"),

@@

This is a shrunken average, with denominator $n+1$. This extra pseudocount has the effect of shrinking the value towards the "neutral" value 0, with the shrinkage becoming stronger if the data is "weak" because the number $|I\cap C_i|$ of positions covered by reads from cell $j$  is low. In the extreme case of none of the reads from cell $j$ covering $I$, the sum becomes 0 and the denominator 1, i.e. $r_{Ij}=0$ in this case.

>>>>

While this simple procedure is straight-forward and produces satisfactory results, the coarse resolution of large genomic tiles and the sparse read coverage of single-cell methylomes still pose significant challenges in the analysis of scBS data.
Here, we propose three improvements that, as we show afterwards, considerably increase the information content of the processed data. We will first explain the proposed improvement in a qualitative manner, while mathematical details are given later in the Methods section. Then, we will demonstrate the improvement using benchmarks.

@@

We also presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms. The package also offers functionality for all other steps of data preprocessing that are needed to get from the output of methylation callers such as Bismark and methylpy to a cell\texttimes region matrix that is suitable as input for standard analysis tools that work in a reduced-dimensionality setting, thus allowing the user to use, e.g., well established tools from scRNA-seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses, etc. By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis.

@@

We write $x_{ij}$ for the methylation status of CpG $i$ in cell $j$. The index $i$ runs over all CpG positions present in the genome, the index $j$ over all cells in the assay. We write $x_{ij}=0$ if position $i$ was found to be unmethylated in cell $j$ by bisulfite sequencing, $x_{ij}=1$ if it was methylated, and $x_{ij}=\text{NA}$ if position $i$ was not covered by reads from cell $j$ and the methylation status is therefore not available (``NA''),

@@

This is a shrunken average, with denominator $n+1$. This extra pseudocount has the effect of shrinking the value towards the ``neutral'' value 0, with the shrinkage becoming stronger if the data is ``weak'' because the number $|I\cap C_i|$ of positions covered by reads from cell $j$  is low. In the extreme case of none of the reads from cell $j$ covering $I$, the sum becomes 0 and the denominator 1, i.e. $r_{Ij}=0$ in this case.
